---
title: "bakerH_Week6_Assignment1"
author: "Heath Baker"
date: "6/18/2021"
output:
  word_document: default
---

```{r,include=FALSE}
library(tidyverse)
library(tidymodels)
```

```{r}
trucks = read_csv("trucks.csv")
```

```{r}
str(trucks)
summary(trucks)
```

### Task 1:  
```{r}
ggplot(trucks, aes(x=Distance, y=Speeding)) + geom_point()
```

Yes - there appears to be at least 2 clusters, between distance 0-100, and above 100. Though possibly 4 natural clusters. 

**The argument for 4:**  
Distance < 100:  
There are two masses. One Thick group at the bottom under speeding values of ~12.5,   
and above speeding values of ~ 12.5  

Distance > 100:  
Speeding below 25 and above 25 seem to be two distinct groupings here.  


### Task 2:  
```{r}
trucks = trucks %>% select(-Driver_ID)
```

```{r}

kmeans_recipe = recipe(~ Distance + Speeding, trucks) 

trucks_dummy = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks_dummy = prep(trucks_dummy, trucks) #prepares the recipe

trucks_cleaned = bake(trucks_dummy, trucks) #applies the recipe and yields a data frame
```

```{r}
#str(trucks_cleaned)
summary(trucks_cleaned)
```
Means are now 0, should now be scaled and centered.  

### Task 3:  

Perform k-means clustering with a pre-specified number of clusters. We use the scaled and dummied data frame.  
```{r}
set.seed(64)
clust1 = kmeans(trucks_cleaned, centers = 2)
#clust
```
Attach cluster to dataset  
```{r}
trucks_clust1 = trucks
```

```{r}
trucks_clust1 = augment(clust1, trucks)
#str(trucks_clust1)
```

Plots  
```{r}
ggplot(trucks_clust1, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point() 
```

Clusters seem to be set to the obvious two groups of Distance > 100, and distance < 100.  
Though there seems to be one outlier in group 1, Distance of around 120 or so.  

### Task 4:  
Perform k-means clustering with a pre-specified number of clusters.
```{r}
set.seed(412)
clust2 = 
  tibble(k = 1:8) %>%
  mutate(
    kclust = map(k, ~kmeans(trucks_cleaned, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks_cleaned)
  )

#clust2
```

```{r}
clusters = 
  clust2 %>%
  unnest(cols = c(tidied))

assignments = 
  clust2 %>% 
  unnest(cols = c(augmented))

clusterings = 
  clust2 %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```
   
I would still argue that 4 clusters is the best type of grouping for this dataset.  
They seem to be distinct groups, and have more specified area than just the two obvious groups.  
   
### Task 5
```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

It appears that, to be "within the elbow", k=4 seems to be the best value of k in this dataset.

### Task 6:  
```{r}
set.seed(64)
clust_final = kmeans(trucks_cleaned, centers = 4)
```

```{r}
trucks = augment(clust_final, trucks)
#str(trucks)
```

Plots  
```{r}
ggplot(trucks, aes(x=Distance,y=Speeding,color=factor(.cluster))) + geom_point() 
```

Though the cutoff points arent' exactly as I had predicted, the 4 groups are fairly close.  
Obvious separations still exist at Distance 125.   

Under 125:   
There is a cluster barrier around Speeding 20,  

Over 125:  
There is a cluster barrier just around Speeding 38.  
