---
title: "Random Forests Assignment"
author: "Heath Baker"
date: "6/7/2021"
output: word_document
---

```{r, include=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(gridExtra)
library(vip)
library(ranger)
```

Loading Data:  
```{r}
drug_data = read_csv("drug_data.csv")
```

```{r}
names(drug_data) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity","Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive","SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis","Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh","LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug_data)
```

### Data Definitions:  
- Nscore - Neuroticism  
- Escore - Extraversion  
- Oscore - Openness to experience  
- Ascore - Agreeableness  
- Cscore - Conscientiousness  
- SS - Sensation Seeking measured by ImpSS  

```{r}
drug_data[drug_data == "CL0"] = "No"
drug_data[drug_data == "CL1"] = "No"
drug_data[drug_data == "CL2"] = "Yes"
drug_data[drug_data == "CL3"] = "Yes"
drug_data[drug_data == "CL4"] = "Yes"
drug_data[drug_data == "CL5"] = "Yes"
drug_data[drug_data == "CL6"] = "Yes"
```

```{r}
drug_clean = drug_data %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44","45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =c("Under16", "At16", "At17", "At18", "SomeCollege","ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
  mutate(Country = factor(Country,labels = c("USA", "NewZealand", "Other", "Australia","Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,labels = c("Black", "Asian", "White", "White/Black", "Other","White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
  select(-ID)
```

Focus on Nicotine use:
```{r}
drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
#names(drug_clean)
```

### Task 1: Missing Data Check
```{r}
summary(drug_clean)
```
Found no missing data in this data, continuing with assignment

### Task 2: Split
```{r}
set.seed(1234) 
drug_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) #70% in training
train = training(drug_split)
test = testing(drug_split)
```

### Task 3: Visualizations
```{r}
#str(train)
```

Factors as bar plots, Numerics as boxplots:  

```{r}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))

grid.arrange(p1,p2,p3,p4)
```

Age - Seems to be a good indicator of nicotine use, the higher the age -> less likely to use Nicotine  
Gender - Seems like males are more likely to use nicotine  
Education- It seems like different education levels correspond with likelihood of nicotine use, a decent indicator  
Country - It looks like different countries have varying rates of Nicotine usage, could be a good indicator.  


```{r}
p4 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 90))

p1 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4, ncol = 2)
```

Nscore - Perhaps higher Neuroticism indicates likelihood of Nicotine usage, though not a large margin.  
Escore - Extraversion does not look like a good indicator of Nicotine usage  
Oscore - Openness to Experience seems to be a good indicator  
Ethnicity - Some ethnicities lookt o be better indicators than others on Nicotine usage  

```{r}
p1 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p2 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p3 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p4 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4, ncol = 2)
```

Ascore - Agreeableness - Looks like those that are less agreeable are more likely to use Nicotine  
Cscore - Conscientiousness - Looks like those those that are less conscientious are more likely to use Nicotine  
Impulsive - It looks like those with higher impulsiveness are more likely to use Nicotine  
SS - "Sensation Seeing", those marked "yes" are more likely to use Nicotine  

### Task 4: Random Forest Model
Set up our folds for cross-validation  
```{r}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)
```

Random forest with an R-defined tuning grid 
```{r}
drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% #add tuning of mtry and min_n parameters
  #setting trees to 100 here should also speed things up a bit, but more trees might be better
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

rf_grid = grid_regular(
  mtry(range = c(2, 8)), #these values determined through significant trial and error
  min_n(range = c(5, 20)), #these values determined through significant trial and error
  levels = 10
)

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

set.seed(123)
rf_res = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid #use the tuning grid
)
```

```{r}
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

No shape forming, looks like good pretty good perfornance

```{r}
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```

Looks like best accuracy is Min_N of 6, and mtry of 16

### Task 5: Finalize Workflow:
```{r}
best_rf = select_best(rf_res, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
```
Best Mtry =   6  
Best Min_n = 16  
(confirmed previous conjecture in select_best)  

```{r}
#fit the finalized workflow to our training data
final_rf_fit = fit(final_rf, train)
```

Variable importance
```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

The most important Variable is SS - "Sensation Seeking", followed by Country UK, age group 45-54, then conscientiousness and openness to experience indices. 

### Task 6: Model Performance

Predictions:
**Training Set**
```{r}
trainpredrf = predict(final_rf_fit, train)
#head(trainpredrf)
```

Confusion matrix
```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```
Accuracy on Training set is 91.96%. Seems like a good model from here, a great improvement over the naive model at 67.07%

**Testing Set
```{r}
testpredrf = predict(final_rf_fit, test)
#head(testpredrf)
```

Confusion matrix
```{r}
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

Accuracy on Testing set is 70.19%. This gives me pause as an increase of accuracy of only, 3.17% from the Naive model is concerning.  
(as well as a 21.77% decrease from the training set)  

### Task 7: Real-World
This model could still be helpful in the real world. Accuracy for these measures might not necessarily be the goal here. Being used as predictors for such behaviors such as drug addiction, in real-world scenarios it could be helpful in healthcare situations to use cautious models. The caution exercised here could help to identify individuals that may become addicted to certain substances, and have them on a mailing list, or a sort of "watchlist" for addictive substance abuse - in order to be more proactive and prevent individuals from abusing differrent substances (as they are also identified in this dataset).

```{r}
#saveRDS(final_rf_fit, "final_rf_fit-assignment2.rds")
```
 